---
title: "Homework 8"
author: "Paul Perez"
date: "4/25/2021"
output:
  html_document: default
  pdf_document: default
subtitle: 'Applied Predictive Modeling - Non Linear Regression Models'
---


```{r}
library(tidyverse)
library(e1071)
library(forecast)
library(AppliedPredictiveModeling)
library(mlbench)
library(nnet)
library(neural)
library(RSNNS)
library(plotmo)
library(caret)
```

# Exercise 7.2
Friedman (1991) introduced several benchmark data sets create by simulation. One of these simulations used the following nonlinear equation to create data: 

$$
y = 10 sin (\pi x_1 x_2) + 20(x_3−0.5)^2 + 10x_4 +5x_5 + N(0,\sigma^2)
$$

where the x values are random variables uniformly distributed between [0, 1] (there are also 5 other non-informative variables also created in the simulation). The package mlbench contains a function called `mlbench.friedman1` that simulates these data:

```{r}
library(mlbench)
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will five the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
featurePlot(trainingData$x, trainingData$y)
```

```{r}
## or other methods.

## This creates a list with a vector 'y' and a matrix
## of predictors 'x'.  Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

### KNN

```{r}
require(caret)
knnModel <- train(x = trainingData$x, 
                  y = trainingData$y,
                  method = "knn",
                  preProcess = c("center", "scale"),
                  tuneLength = 10)
knnModel
```

```{r}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## performance values
postResample(pred = knnPred, obs = testData$y)
```

```{r}
knnModel$bestTune
```


```{r}
ggplot(knnModel)
```

```{r}
knnModel$finalModel
```


```{r}
varImp(knnModel)
plot(varImp(knnModel))
```

We see 9 important variables with the Knn model.

\newpage

### MARS 
```{r}
library(plotmo)
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:38)
# Fix the seed tso that the results can be reproduced
set.seed(123)
marsTuned <- train(x = trainingData$x, 
                  y = trainingData$y,
                  method = "earth",
                  tuneGrid = marsGrid,
                  preProcess = c("center", "scale"),
                  tuneLength = 10)
marsTuned
```

```{r}
marsTuned$bestTune
```


```{r}
ggplot(marsTuned)
```

We can see the optimal model here has 15 terms.

```{r}
marsTuned$finalModel
```


```{r}
varImp(marsTuned)
plot(varImp(marsTuned))
```

We can see that with the Mars Model, there is are 5 variables of importance.

\newpage

### Neural Networks
```{r}
## Create a specific candidate set of models to evaluate: 
nnetGrid <- expand.grid(decay = c(0, 0.01, .1), size = c(1:10), bag = FALSE) 
set.seed(123) 
nnetTuned <- train(trainingData$x, trainingData$y,  
                  method = "avNNet",  
                  tuneGrid = nnetGrid,  
                  trControl = trainControl(method = "cv", number = 10),
                  preProc = c("center", "scale"),  
                  linout = TRUE,  trace = FALSE,  
                  MaxNWts = 10 * (ncol(trainingData$x) + 1) + 10 + 1, 
                  maxit = 500)
nnetTuned
```

```{r}
nnetTuned$bestTune
```

```{r}
nnetTuned$finalModel
```

```{r}
ggplot(nnetTuned)
```

```{r}
varImp(nnetTuned)
plot(varImp(nnetTuned))
```

With the neural network model, we see that there are 9 variables of importance.

\newpage

### Support Vector Machines

```{r}
set.seed(123)
svmTuned <- train(trainingData$x, trainingData$y, 
                     method = "svmRadial", 
                     preProc = c("center", "scale"),
                     tuneLength = 14, 
                     trControl = trainControl(method = "cv"))
svmTuned
```

```{r}
svmTuned$bestTune
```


```{r}
svmTuned$finalModel
```


```{r}
ggplot(svmTuned)
```


```{r}
varImp(svmTuned)
plot(varImp(svmTuned))
```

With out support vector model, we see there are 9 variables of importance.

#### Comparison of Models

```{r}
marsPred <- predict(marsTuned, newdata = testData$x)
marsPv <- postResample(pred = marsPred, obs = testData$y)
marsPv

nnPred <- predict(nnetTuned, newdata = testData$x)
nnPv <- postResample(pred = nnPred, obs = testData$y) 
nnPv

svmPred <- predict(svmTuned, newdata = testData$x)
svmPv <- postResample(pred = svmPred, obs = testData$y) 
svmPv

knnPred <- predict(knnModel, newdata = testData$x)
knnPv <- postResample(pred = knnPred, obs = testData$y)
knnPv
```

We can see that with our models, the MARS model presents the greatest r-Squared and lowest RMSE. Looking at the variables of importance for the MARS model, we see those variables named X1 - X5.

\newpage

# Exercise 7.5

Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several nonlinear regression models.

For week 10's homework. I created a ridge regression model. Below is the code. Following that, are the non-linear models for this exercise.

### (a) Which nonlinear regression model gives the optimal resampling and test set performance?

```{r}
library(RANN)
data("ChemicalManufacturingProcess")
preprocessing <- preProcess(ChemicalManufacturingProcess[,-1], 
                            method = c("center", "scale", "knnImpute", "corr", "nzv")) 
preprocess <- predict(preprocessing, ChemicalManufacturingProcess[,-1])
```

```{r}
yield <- as.matrix(ChemicalManufacturingProcess$Yield)
set.seed(123)
split2 <- createDataPartition(yield, p = 0.8, list = FALSE, times = 1)

Xtrain.data2  <- preprocess[split2, ]
xtest.data2 <- preprocess[-split2, ]
Ytrain.data2  <- yield[split2, ]
ytest.data2 <- yield[-split2, ]
```

```{r}
ridge.grid <- data.frame(.lambda = seq(0, .1, length = 15)) 
set.seed(123) 
ctrl <- trainControl(method = "cv", number = 10)
ridge.mod2 <- train(x=Xtrain.data2, 
                    y=Ytrain.data2, 
                    method="ridge", 
                    tuneGrid=ridge.grid, 
                    trControl= ctrl)
ridge.mod2
plot(ridge.mod2)
```


```{r}
ridge.mod2$bestTune
```

```{r}
ridge.mod2$finalModel
```


```{r}
ggplot(ridge.mod2)
```


```{r}
varImp(ridge.mod2)
plot(varImp(ridge.mod2))
```

\newpage

### MARS

```{r}
marsGrid <- expand.grid(.degree = 1:3, .nprune = 2:100)
set.seed(123)
chemMarsTuned <- train(Xtrain.data2, Ytrain.data2,
                   method = "earth",
                   tuneGrid = marsGrid,
                   trControl = trainControl(method = "cv", number = 10))
chemMarsTuned
```

```{r}
chemMarsTuned$bestTune
```

```{r}
chemMarsTuned$finalModel
```


```{r}
ggplot(chemMarsTuned)
```


```{r}
varImp(chemMarsTuned)
plot(varImp(chemMarsTuned))
```


### Support Vector Machine

```{r}
set.seed(123)
chemSvmTuned <- train(Xtrain.data2, Ytrain.data2,
                     method = "svmRadial",
                     preProc = c("center", "scale"),
                     tuneLength = 14,
                     trControl = trainControl(method = "cv"))
chemSvmTuned
```

```{r}
chemSvmTuned$bestTune
```

```{r}
chemSvmTuned$finalModel
```


```{r}
ggplot(chemSvmTuned)
```


```{r}
varImp(chemSvmTuned)
plot(varImp(chemSvmTuned))
```


### KNN

```{r}
set.seed(123)
chemKnnTuned <- train(Xtrain.data2,
                        Ytrain.data2,
                        method = "knn",
                        tuneGrid = data.frame(.k = 1:20),
                        trControl = trainControl(method = "cv"))
chemKnnTuned
```

```{r}
chemKnnTuned$bestTune
```

```{r}
chemKnnTuned$finalModel
```


```{r}
ggplot(chemKnnTuned)
```


```{r}
varImp(chemKnnTuned)
plot(varImp(chemKnnTuned))
```

#### Comparison of Models

```{r}
ridgePred <- predict(ridge.mod2, newdata = xtest.data2)
ridgePv2 <- postResample(pred = ridgePred, obs = ytest.data2)
ridgePv2

marsPred2 <- predict(chemMarsTuned, newdata = xtest.data2)
marsPv2 <- postResample(pred = marsPred2, obs = ytest.data2)
marsPv2

svmPred2 <- predict(chemSvmTuned, newdata = xtest.data2)
svmPv2 <- postResample(pred = svmPred2, obs = ytest.data2) 
svmPv2

knnPred2 <- predict(chemKnnTuned, newdata = xtest.data2)
knnPv2 <- postResample(pred = knnPred2, obs = ytest.data2)
knnPv2
```

The SVM model shows strong results having a 1.230 RMSE and an r-squared of 0.549.
\newpage

### (b) Which predictors are most important in the optimal nonlinear regression model?

### MARS

```{r}
varImp(chemSvmTuned)
```

### Ridge
```{r}
varImp(ridge.mod2)
```

The outputs for importance of variables with both models show the same variables.

\newpage

### (c) Explore the relationships between the top predictors and the response

```{r}
cor(yield, ChemicalManufacturingProcess$ManufacturingProcess32)
cor(yield, ChemicalManufacturingProcess$ManufacturingProcess13)
```

Looking at the top two Manufacturing Process variables, we  can see a positive correlation with Process 32 and a negative correlation with Process 13.
