---
title: "Homework 7"
author: "Paul Perez"
date: "4/18/2021"
output:
  html_document: default
  pdf_document: default
subtitle: 'Applied Predictive Modeling - Linear Regression and Its Cousins'
---


# Exercise 6.2
Developing a model to predict permeability (see Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:

## (a) Start R and use these commands to load the data:

```{r, warning=FALSE, results='hide', echo=FALSE, message=FALSE}
library(AppliedPredictiveModeling)
data(permeability)
```

The matrix `fingerprints` contains the 1,107 binary molecular predictors for the 165 compounds, while permeability contains permeability response.

## (b) The fingerprint predictors indicate the presence or absence of substructures of a molecule and are often sparese meaning that relatively few of the molecules contain each substructure. Filter out the predictors that have low frequencies using `nearZeroVar` function from the caret package. How many predictors are left for modeling?

```{r}
library(caret)
nzv <- nearZeroVar(fingerprints)
not.nzv <- fingerprints[, -nzv]
ncol(not.nzv)
```

## (c) Split the data into a training and test set, pre-process the data, and tune a PLS model. How many latent variables are optimal and what is the corresponding resampled estimate of R^2?

```{r}
set.seed(123)
split <- createDataPartition(permeability, p = 0.8, list = FALSE, times = 1)

Xtrain.data  <- not.nzv[split, ] #fingerprints train
xtest.data <- not.nzv[-split, ] #fingerprints test
Ytrain.data  <- permeability[split, ] #permability train
ytest.data <- permeability[-split, ] #permability test
```

```{r}
ctrl <- trainControl(method = "cv", number = 10)
pls.mod <- train(x = Xtrain.data, 
                y = Ytrain.data, method = "pls", 
                tuneLength = 20, 
                trControl = ctrl, 
                preProc = c("center", "scale"))
pls.mod
plot(pls.mod)
```


```{r}
pls.mod$bestTune
summary(pls.mod$finalModel)
```

6 is the best tuning parameter to minimze cross validation error. 

## (d) Predict the response for the test set. What is the test set estimate of R^2?

```{r}
predictions <- predict(pls.mod, xtest.data)
cbind(RMSE = RMSE(predictions, ytest.data),  R.squared = caret::R2(predictions, ytest.data))
```

```{r}
plot(predictions, 
     col = "darkblue", 
     main = "Observed vs. Predicted - Partial Least Squares Regression Model",
     xlab = "", 
     ylab = "Predictions")

par(new = TRUE)

plot(ytest.data, 
     col = "orange", 
     axes=F, 
     ylab = "", 
     xlab="Observed") 

abline(0, 1, col='green')
```

\newpage

## (e) Try building other models discussed in this chapter. Do any have better predictive performance?

We can perform a similar predition with both Ridge and Lasso regression models.

### Ridge Model

```{r}
set.seed(123)
ridge.mod <- train(x=Xtrain.data,
                  y=Ytrain.data,
                  method='ridge',
                  metric='Rsquared',
                  tuneGrid=data.frame(.lambda = seq(0, 1, by=0.1)),
                  trControl=trainControl(method='cv'),
                  preProcess=c('center','scale')
                  )
ridge.mod
plot(ridge.mod)
```

```{r}
ridge.mod$bestTune
summary(ridge.mod$finalModel)
```

```{r}
predictions <- predict(ridge.mod, xtest.data)
cbind(RMSE = RMSE(predictions, ytest.data),  R.squared = caret::R2(predictions, ytest.data))
```

```{r}
plot(predictions, 
     col = "darkblue", 
     main = "Observed vs. Predicted - Ridge Regression Model",
     xlab = "", 
     ylab = "Predictions")

par(new = TRUE)

plot(ytest.data, 
     col = "orange", 
     axes=F, 
     ylab = "", 
     xlab="Observed") 

abline(0, 1, col='green')
```

\newpage

### Lasso Model

```{r}
set.seed(123)
lasso.mod <- train(x=Xtrain.data,
                  y=Ytrain.data,
                  method='lasso',
                  metric='Rsquared',
                  tuneGrid=data.frame(.fraction = seq(0, 0.5, by=0.05)),
                  trControl=trainControl(method='cv'),
                  preProcess=c('center','scale')
                  )
lasso.mod
plot(lasso.mod)
```

The model has some issues fitting as it failes for `Fold09`. 

```{r}
lasso.mod$bestTune
summary(lasso.mod$finalModel)
```

```{r}
predictions <- predict(lasso.mod, xtest.data)
cbind(RMSE = RMSE(predictions, ytest.data), R.squared = caret::R2(predictions, ytest.data))
```

```{r}
plot(predictions, 
     col = "darkblue", 
     main = "Observed vs. Predicted - Lasso Regression Model",
     xlab = "", 
     ylab = "Predictions")

par(new = TRUE)

plot(ytest.data, 
     col = "orange", 
     axes=F, 
     ylab = "", 
     xlab="Observed") 

abline(0, 1, col='green')
```

## (f) Would you recommend any of your models to replace the permeability laboratory experiment?

While the three model's had similar R^2 values, the Ridge regression model had a greater RMSE and would be my recommendation to replace the permability laboratory experiment.

\newpage

# Exercise 6.3
A chemical manufacturing process for a pharmaceutical product was discussed in Sect. 1.4. In this problem, the objective is to understand the relationship between biological measurements of the raw materials (predictors), measurements of the manufacturing process (predictors), and the response of product yield. Biological predictors cannot be changed but can be used to assess the quality of the raw material before processing. On the other hand, manufacturing process predictors can be changed in the manufacturing process. Improving product yield by 1% will boost revenue by approximately one hundred thousand dollars per batch:

## (a) Start R and use these commands to load the data:

```{r}
library(AppliedPredictiveModeling)
data("ChemicalManufacturingProcess")
```

The matrix processPredictors contains the 57 predictors (12 describing the input biological material and 45 describing the process predictors) for the 176 manufacturing runs. `yield` contains the percent yield for each run.

## (b) A small percentage of cells in the predictor set contain missing values. Use an imputation function to fill in the missing values (e.g., See Sect. 3.8)

```{r}
library(RANN)
preprocessing <- preProcess(ChemicalManufacturingProcess[,-1], 
                            method = c("center", "scale", "knnImpute", "corr", "nzv")) 

preprocess <- predict(preprocessing, ChemicalManufacturingProcess[,-1])
```

## (c) Split the data into a training and a test set, pre-process the data, and tune am odel of your choice from this chapter. What is the optimal value of the performance metric?

```{r}
yield <- as.matrix(ChemicalManufacturingProcess$Yield)
set.seed(123)
split2 <- createDataPartition(yield, p = 0.8, list = FALSE, times = 1)

Xtrain.data2  <- preprocess[split2, ]
xtest.data2 <- preprocess[-split2, ]
Ytrain.data2  <- yield[split2, ]
ytest.data2 <- yield[-split2, ]
```

```{r}
ridge.grid <- data.frame(.lambda = seq(0, .1, length = 15)) 
set.seed(123) 
ridge.mod2 <- train(x=Xtrain.data2, 
                    y=Ytrain.data2, 
                    method="ridge", 
                    tuneGrid=ridge.grid, 
                    trControl= ctrl)
ridge.mod2
plot(ridge.mod2)
```


```{r}
ridge.mod2$bestTune
```

We can see in the output, that our optimal value is a lambda of 0.1.

## (d) Predict the response for the test set. What is the value of the performance metric and how does this compare with the resampled performance metric of the training set?

```{r}
predictions <- predict(ridge.mod2, xtest.data2)
cbind(RMSE = RMSE(predictions, ytest.data2),  R.squared = caret::R2(predictions, ytest.data2))
```

The response of the test set performed better than the resampled metrics of the training set.

## (e) Which predictors are most important in the model you have trained? Do either the biological or process predictors dominate the list?

```{r, fig.height= 10, fig.width= 9}
varImp(ridge.mod2)
plot(varImp(ridge.mod2))
```

We can see from the chart above that the process predictors dominate the list.

## (f) Explore the relationships between each of the top predictors and the response. How could this information be helpful in improving yield in the future runs of the manufacturing process?

```{r}
cor(yield, ChemicalManufacturingProcess$ManufacturingProcess13)
cor(yield, ChemicalManufacturingProcess$BiologicalMaterial06)
```

We can see that Biological materials are used to asses the quality of raw materials before processing, and with positibe results, the yield of the product may increase. Looking at the top Biological material, we can see the positive but moderately correlation to the response variable.

Considering the above statement, seeing a negative correlation with the manufacturing process, we understand that the product may not come out good.