---
title: "Homework 1"
author: "Paul Perez"
date: "2/14/2021"
output:
  pdf_document: default
  html_document: default
subtitle: 'Forecasting: Principles and Practice - Time Series Graphics'
---

```{r, warning=FALSE}
library(fpp3)
library(ggplot2)
```

# Exercise 2.10 - 1
Use the help function to explore what the series `gafa_stock`, `PBS`, `vic_elec` and `pelt` represent.

## a.
Use `autoplot()` to plot some of the series in these data sets.

## b.
What is the time interval of each series?

### gafa_stock
The data in `gafa_stock` represents historical stock prices ($USD) for Google, Amazon, Facebook, and Apple from 2014 - 2018.

```{r}
gafa_stock
```

Looking at the `tsibble` object above, we have 5,032 records with 8 values per record. Those are `Symbol`, `Date`, `Open`, `High`, `Low`, `Adj_Close`, and `Volume`. The data [1D] shows that the data is at a daily level, and the Symbol [4] shows that we have four tickers.

```{r}
gafa_stock %>% distinct(Symbol)
```

```{r}
autoplot(gafa_stock)
```

Using `autoplot()` without specifying a variable to plot, we can see the four stock tickers plotted along the same x-axis using the `Open` variable which is the stocks opening price. Below, we can specifically plot the `Close` price by adding it as the second argument in the `autoplot()` function.

```{r}
autoplot(gafa_stock, Close)
```

### PBS
The data in '`PBS` represents monthly medicare prescription data in Australia.
```{r}
PBS
```

Looking at the `tsibble` object above, we have 65,219 records with 9 values per record. Those are `Month`, `Concession`, `Type`, `ATC1`, `ATC1_desc`, `ATC2`, `ATC2_desc`, `Scripts`, `Cost`. The data [1M] shows that the data is at a monthly level, and the Concession, Type, ATC1, and ATC2 [336] shows that we have each of those values have 336 unique values.

Since there are various levels in this nested `tsibble` object, we'll make sure to filter and aggregate this data accordingly.

```{r}
PBS %>% distinct(ATC2)
```

```{r}
PBS %>%
  filter(ATC2 == "A01") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) -> a01


PBS %>%
  filter(ATC2 == "A02") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) -> a02
```

```{r}
autoplot(a01, TotalC)
```

```{r}
autoplot(a02, TotalC)
```


### vic_elec
The data in `vic_elec` represents half-hourly electicity demand for Victoria, Australia given three values; Demand, Temperature, and Holiday.

```{r}
vic_elec
```

Looking at the `tsibble` object above, we have 52,608 records with 5 values per record. Those are `Time`, `Demand`, `Temperature`, `Date`, and `Holiday`. The data [30m] shows that the data is at a half-hour time period.

```{r}
autoplot(vic_elec, Demand)
```

```{r}
autoplot(vic_elec, Temperature)
```

### pelt
The data in `pelt` represents Hudson Bay Company trading records for Snowshow Hare and Canadian Lynx furs from 1845 to 1935.
```{r}
autoplot(pelt)
```

Looking at the `tsibble` object above, we have 52,608 records with 5 values per record. Those are `Time`, `Demand`, `Temperature`, `Date`, and `Holiday`. The data [30m] shows that the data is at a half-hour time period.


\newpage
# Exercise 2.10 - 2
Use `filer()` to find what days corresponded to the peak closing price for each of the four stocks in `gafa_stock`.

\newpage
# Exercise 2.10 - 3
Download the file `tute1.csv` from the book website, open in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly  sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation. 

## a.
You can read the data into R with the following script:

```{r}
tute1 <- readr::read_csv("tute1.csv")
View(tute1)
```

## b. 
Convert the data to time series

```{r}
mytimeseries <- tute1 %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)
```

## c.
Construct time series plots of each of the three series
```{r}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) + 
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```
Check what happens when you don't include `facet_grid()`.


\newpage
# Exercise 2.10 - 6
Create time plots of the following four time series: `Bricks` from `aus_production`, `Lynx` from `pelt`, `Close` from `gafa_stock`, `Demand` from `vic_elect`.

- Use `?` (or `help()`) to find out about the data in each series.
- For the last plot_modify the axis labels and title.



